\section{Active learning (Stage 1)}
\label{sec:visualizer:al}

\subsection{Decision tree classification of user interests}
\label{sec:visualizer:al:tree}

A decision tree is a method of classifying and labeling plots. An ``active
learner'' adapts as the process moves forward, choosing its points of query
intelligently. Active learning increases efficiency when searching through the
hypothesis space, which is any fitted tree that agrees with the labeled data as
much as possible~\cite{dasgupta2011}. Every time new data (a new label) is
received, the hypothesis space shrinks as the label removes certain classifiers
from the running~\cite{dasgupta2011}. An active learner queries from ambiguous
parts of the current hypothesis space so as to shrink it as quickly as possible.
It is problematic to start from scratch, however; how does the system determine
the best first point of ambiguity when it knows nothing (the hypothesis space is
everything)? While this problem is difficult, we can exploit the fact that the
user is already providing a numerical model that they believe to be a good
representation of the data which they would like the visualization system to
check visually. Given this data, the system builds a decision tree that utilizes
the various properties of the plots to determine whether one is interesting or
uninteresting. Doing so greatly narrows the hypothesis space and makes it easier
to determine points of ambiguity. However, to reconcile with the fact that the
user wishes to check the numerical model and may not necessarily believe it is a
good representation of fit, the learner must then perform several line-up tests
(Section ~\ref{sec:visualizer:al:lineup}) to check whether the initial decision
tree is a proper fit. As the user then proceeds to label various conditional
plots as ``interesting'' or ``not interesting,'' the classifier learns the
user’s interest and continues to evolve. This models plot characteristics that
the user found interesting to study.

\subsection{Selective plot generation}
\label{sec:visualizer:al:alplotgeneration}

((To build a better classifier, we want to have the user label plots that the
system (at the time) is uncertain about. This is where active learning comes in
since we want to build our system to cleverly give users vital plots to label so
the system can best learn the user’s interests. The system will have to
determine which features it is uncertain about classifying and return a plot
matching those characteristics to the user. ))
