\section{Healthcare stock data}
\label{sec:usage:data}

The dataset contains 4110 observations of 43 stock prices from the healthcare 
industry observed from 1998 to 2014. The data is cleaned to get rid of the 
dependency on time and yearly seasonality so that each observation of a stock 
may better simulate an independent draw from some distribution. This is done 
with the usage of the \texttt{sstl} function from the \texttt{Rsafd} package; 
after fitting the seasonal and trend components, the residuals are used as the 
new dataset. A list of ticker names and the associated firm may be found below

\tablespacing
\begin{longtable}{p{0.15\linewidth} p{0.5\linewidth}}
	
	% First page heading
	\caption[Healthcare firms and stock tickers.]{Healthcare firms and stock 
	tickers.} 
	\label{tab:usage:tickers}\\
	\toprule
	\textbf{Ticker} & \textbf{Firm} \\
	\midrule
	\endfirsthead
	
	% Future page heading
	\caption[]{(continued)}\\
	\toprule
	\textbf{Ticker} & \textbf{Firm} \\
	\midrule
	\endhead
	
	% Page footer
	\midrule
	\multicolumn{2}{r}{(Continued on next page)}\\
	\endfoot
	
	% Last page footer
	\bottomrule
	\endlastfoot
	
	ABT & ABBOTT LABORATORIES \\
	ABBV & ABBVIE INC \\
	ACT & ACTAVIS PLC \\
	AET & AETNA INC \\
	A & AGILENT TECHNOLOGIES INC \\
	ALXN & ALEXION PHARMACEUTICALS INC \\
	AGN & ALLERGAN INC \\
	ABC & AMERISOURCEBERGEN CORP \\
	AMGN & AMGEN INC \\
	BAX & BAXTER INTERNATIONAL INC \\
	BDX & BECTON DICKINSON AND CO \\
	BIIB & BIOGEN IDEC INC \\
	BSX & BOSTON SCIENTIFIC CORP \\
	BMY & BRISTOL-MYERS SQUIBB CO \\
	CAH & CARDINAL HEALTH INC \\
	CFN & CAREFUSION CORP \\
	CELG & CELGENE CORP \\
	CERN & CERNER CORP \\
	CI & CIGNA CORP \\
	COV & COVIDIEN PLC \\
	BCR & CR BARD INC \\
	DVA & DAVITA HEALTHCARE PARTNERS I \\
	XRAY & DENTSPLY INTERNATIONAL INC \\
	EW & EDWARDS LIFESCIENCES CORP \\
	LLY & ELI LILLY \& CO \\
	ESRX & EXPRESS SCRIPTS HOLDING CO \\
	FRX & FOREST LABORATORIES INC \\
	GILD & GILEAD SCIENCES INC \\
	HSP & HOSPIRA INC \\
	HUM & HUMANA INC \\
	ISRG & INTUITIVE SURGICAL INC \\
	JNJ & JOHNSON \& JOHNSON \\
	LH & LABORATORY CRP OF AMER HLDGS \\
	MCK & MCKESSON CORP \\
	MDT & MEDTRONIC INC \\
	MRK & MERCK \& CO. INC. \\
	MYL & MYLAN INC \\
	PDCO & PATTERSON COS INC \\
	PKI & PERKINELMER INC \\
	PRGO & PERRIGO CO PLC \\
	PFE & PFIZER INC \\
	DGX & QUEST DIAGNOSTICS INC \\
	REGN & REGENERON PHARMACEUTICALS \\
	STJ & ST JUDE MEDICAL INC \\
	SYK & STRYKER CORP \\
	THC & TENET HEALTHCARE CORP \\
	TMO & THERMO FISHER SCIENTIFIC INC \\
	UNH & UNITEDHEALTH GROUP INC \\
	VAR & VARIAN MEDICAL SYSTEMS INC \\
	VRTX & VERTEX PHARMACEUTICALS INC \\
	WAT & WATERS CORP \\
	WLP & WELLPOINT INC \\
	ZMH & ZIMMER HOLDINGS INC \\
	ZTS & ZOETIS INC \\
	
\end{longtable}
\bodyspacing

The data has been split into testing and training sets of equal length (2050 
observations) between 1998 and 2014. The idea is that we are standing there in 
the middle (at 3/13/2006 to be precise) and wish to select our portfolio of 
stocks on that day. Although we are uncertain of the future, we do have the 
historical price performance of each and may also extract qualities such as 
drift, volatility, returns, etc. from the data. As such, correlation graphs 
(see Section~\ref{sec:intro:correlation}) of the data may be computed and used 
to inform the stock selection. Once the portfolio has been 
selected and purchased, the ``buy and hold'' model is implemented, and we 
observe the yearly returns of the same portfolio until 2014. The next section  
details the stock selection procedure given a correlation graph.
